Marlowe: financial contracts on blockchain\thanks{This work is part of the Cardano project and is supported by IOHK,~\url{https://iohk.io}}

### 2018-07-26 tex/paper.tex
Assded reference to split contracts, and discussion in related work. Changed the title of section 5 and 5.1, and added more about design rationale to the introduction. Added new introduction to the Meadow section. Modified the discussion of "embedded" in contributions.

We make the following contributions in this paper.
\begin{itemize}
\item
Designing a DSL for financial contracts on blockchains: Marlowe.
\item
Defining an executable, small-step semantics of Marlowe in Haskell.
\item 
Embedding of Marlowe in Haskell to extend its expressibility; that is: Marlowe is defined as a Haskell 
\haskellinline{data} type so that generating Marlowe contracts from Haskell is straightforward.
\item
Developing the Meadow tool that allows users to interact with and simulate the operation of Marlowe contracts 
and embedded Marlowe contracts.
\end{itemize}

Having established this model and its semantics we are able to do a number of other things, which we discuss in the 
paper. We can explore how the language will be implemented on an existing blockchain, such as Cardano, and our model has 
shown us that we will need to consider how the evolution of a contract interacts with the blockchain, with miners in the 
case of a `proof of work'-based chain, and with users participating in contract execution. We can also perform analyses 
of Marlowe contracts, based on its formal semantics. Finally, we can explore the design space for Marlowe, and explore 
potential alternative designs.

@@

The paper begins in Section \ref{section:model} by introducing the Marlowe 
model, including the assumptions made in designing it, the types of the principal functions and a description of the 
DSL as an algebraic type, constructor by constructor. 
%Section \ref{section:example-saving} presents an example of a larger Marlowe contract, for incentivised saving. 
Section \ref{section:example-escrow} revisits the escrow example, and shows 
how it is described using a combination of Marlowe and Haskell constructs: that is, we use Marlowe as an embedded DSL. 
Section \ref{section:tool} introduces Meadow, our tool for visualising and interacting with Marlowe contracts.
Section \ref{sec:abstraction} reflects on the system design, showing how Marlowe can be supported on a variety of 
blockchains, and Section \ref{sec:compilation} explores how Marlowe can be implemented. Section \ref{sec:related} 
surveys related work and Section 
\ref{section:next-steps} enumerates the next steps for the project after drawing some conclusions.

@@

We used Haskell because it is the language in which Marlowe is implemented, but it would be easy to embed Marlowe in any 
other language. It would only be necessary to translate its primitives into a data type in that language. In Meadow we 
use Fay \cite{Fay}, which is a subset of Haskell (see Section~\ref{section:tool}).

@@

One of the objectives of Marlowe is to be accessible to a wider audience
than a general purpose language would. With this aim, and to aid Marlowe's
usage by people that does not know or is not familiar with its syntax,
we have developed Meadow, a web tool that supports the interactive construction,
edition, and simulation of smart-contracts written in Marlowe. The
tool is publicly available in the url: \url{https://input-output-hk.github.io/scdsl/}.
In Figure~\ref{fig:full-screenshot-demo}, we provide a screenshot
of Meadow in the middle of simulating the execution of the ``deposit incentive'' contract available
in the GitHub repository (in the file: \texttt{src/DepositIncentive.hs}) and in the examples section of Meadow (on the 
bottom right part).

@@

The reason we chose Fay instead of Haskell for Meadow is a technical one: we wanted Meadow to be run completely inside 
the browser, because that makes it easy to deploy, since the server is only required to send the page and not to 
run any Marlowe-related computation. There exist compilers from Haskell to JavaScript but most of them are not easy to 
bootstrap into JavaScript. Among other reasons, they often rely on the API of the host OS to read and write files.

Fay compiler, on the other hand, is mostly written in pure Haskell, and it is not hard to compile to JavaScript by 
using GHCJS. Nevertheless, it does not include a type checker, it relies in GHC's type-checker; and it also 
tries to read its \haskellinline{Prelude} from disk. We have worked around these issues by disabling type-checking 
within Meadow embedded editor, and by embedding all the required modules as constant strings in a modified version 
of the Fay compiler.

@@

\section{Implementation on blockchain}
\label{section:towards-implementation}

@@

\subsection{Abstraction}
\label{sec:abstraction}

Marlowe abstracts away from a number of concrete details of how blockchains operate. In particular, it is agnostic 
between UTxO-based systems, such as Bitcoin and Cardano SL, and account-based models such as Ethereum; it can be 
implemented on push or pull-based systems, and can be executed on or off-chain.

@@

>>>>

We make the following contributions in this paper.
\begin{itemize}
\item
Designing a DSL for financial contracts on blockchains: Marlowe.
\item
Defining an executable, small-step semantics of Marlowe in Haskell.
\item 
Making Marlowe an embedded DSL in Haskell. This extends the expressibility of the language, as we can use all the facilities of Haskell in defining Marlowe contracts; we achieve this by defining Marlowe as  a Haskell \haskellinline{data} type.
\item
Developing the Meadow tool that allows users to interact with and simulate the operation of Marlowe contracts 
and embedded Marlowe contracts.
\end{itemize}

Having established this model and its semantics we are able to do a number of other things, which we discuss in the 
paper. We can explore how the language will be implemented on an existing blockchain, such as Cardano, and our model has 
shown us that we will need to consider how the evolution of a contract interacts with the blockchain, with miners in the 
case of a `proof of work'-based chain, and with users participating in contract execution. We can also perform analyses 
of Marlowe contracts, based on its formal semantics.

In designing a language like Marlowe we are constrained by the blockchain domain, but within that we do have a range of choices. For example, should we base the language on a system with accounts, like Ethereum, or a UTxO-based model as used by bitcoin? We examine this and other choices after introducing the language and examples of its use.

@@

The paper begins in Section \ref{section:model} by introducing the Marlowe 
model, including the assumptions made in designing it, the types of the principal functions and a description of the 
DSL as an algebraic type, constructor by constructor. 
%Section \ref{section:example-saving} presents an example of a larger Marlowe contract, for incentivised saving. 
Section \ref{section:example-escrow} revisits the escrow example, and shows 
how it is described using a combination of Marlowe and Haskell constructs: that is, we use Marlowe as an embedded DSL. 
Section \ref{section:tool} introduces Meadow, our tool for visualising and interacting with Marlowe contracts.
Section \ref{sec:abstraction} reflects on the design rationale for Marlowe, showing how it can be supported on a variety of 
blockchains, and Section \ref{sec:compilation} explores how Marlowe can be implemented. Section \ref{sec:related} 
surveys related work and Section 
\ref{section:next-steps} enumerates the next steps for the project after drawing some conclusions.

@@

We used Haskell because it is the language in which Marlowe is implemented, but it would be easy to embed Marlowe in any 
other language. It would only be necessary to translate its primitives into a data type in that language. In Meadow we 
use Fay \cite{Fay}, a subset of Haskell that we discuss in more detail in Section~\ref{section:tool}).

@@

For Marlowe to be usable in practice, users need to be able to understand how contracts will behave once deployed to the blockchain, but without doing the deployment. We can do that by simulating their behaviour off-chain, interactively stepping through the evaluation of a contract in a browser. We do this in two stages, first transforming an embedded contact (using features of Haskell) to a pure Marlowe contract, and then interactively stepping through that contract.

To achieve this, and to aid Marlowe's take-up
usage by people that does not know or are not familiar with its syntax,
we have developed Meadow, a web tool that supports the interactive construction,
revision, and simulation of smart-contracts written in Marlowe. The
tool is publicly available in the url: \url{https://input-output-hk.github.io/scdsl/}.
In Figure~\ref{fig:full-screenshot-demo}, we provide a screenshot
of Meadow in the middle of simulating the execution of the ``deposit incentive'' contract available
in the GitHub repository (in the file: \texttt{src/DepositIncentive.hs}) and in the examples section of Meadow (on the 
bottom right part).

@@

The reason that we chose Fay instead of Haskell for Meadow is a technical one: we wanted Meadow to be run completely inside 
the browser, because that makes it easy to deploy, since the server is only required to send the page and not to 
run any Marlowe-related computation. There exist compilers from Haskell to JavaScript but most of them are not easy to 
bootstrap into JavaScript. Among other reasons, they often rely on the API of the host OS to read and write files.

The Fay compiler, on the other hand, is mostly written in pure Haskell, and it is not hard to compile to JavaScript by 
using GHCJS. Nevertheless, it does not include a type checker, it relies in GHC's type-checker; and it also 
tries to read its \haskellinline{Prelude} from disk. We have worked around these issues by disabling type-checking 
within Meadow embedded editor, and by embedding all the required modules as constant strings in a modified version 
of the Fay compiler.

@@

\section{Design and Implementation}
\label{section:towards-implementation}

@@

\subsection{Design Rationale}
\label{sec:abstraction}
\label{sec:design-rationale}

Marlowe abstracts away from a number of concrete details of how blockchains operate. In particular, it is agnostic 
between UTxO-based systems, such as Bitcoin and Cardano SL, and account-based models such as Ethereum; it can be 
implemented on ``push'' or ``pull''-based systems, and can be executed on or off-chain.

@@

\paragraph{Split contracts}
An early work on smart contracts raises the issue that practical contractual situations may well not be amenable to complete formalisation, hence giving rise to a \emph{split}~\cite{split-contracts} between an automated part and a non-automated part that mediates, for example, real assets. We can foresee that this approach will be needed for Marlowe too, if it achieves read-world adoption.
